---
title: "p8105_hw2_fx2212"
author: "Fenglin Xie"
date: "2025-09-29"
output: github_document
---
```{r}
library(tidyverse)
```

## Problem 1 
### clean the data in pols-month.csv

```{r}
pols_month_df = 
  read_csv("fivethirtyeight_datasets/pols-month.csv") |> 
  janitor::clean_names() |> 
  separate(
    col = mon,
    into = c("year", "month", "day")
  ) |> 
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day),
    month = month.name[month],
    president = case_when(
      prez_gop == 1 ~ "gop",
      prez_dem == 1 ~ "dem",
      TRUE ~ NA_character_
    )
  ) |>
  select(-prez_dem, -prez_gop, -day) 
  
```

### clean the data in snp.csv

```{r}
snp_df =
  read_csv("fivethirtyeight_datasets/snp.csv") |> 
  janitor::clean_names() |>
  mutate(
    date = mdy(date)
  ) |> 
  mutate(
    year = year(date),
    month = month(date),
    month = month.name[month]
  ) |> 
  select(year, month, close) |> 
  arrange(year, match(month, month.name))
```

### tidy the unemployment data

```{r}
unemployment_df = 
  read_csv("fivethirtyeight_datasets/unemployment.csv") |> 
  janitor::clean_names() |> 
  pivot_longer(
    cols = jan:dec,
    names_to = "month",
    values_to = "unemployment_rate"
  ) |> 
  mutate(
      month = case_when(
      month == "jan" ~ "January",
      month == "feb" ~ "February", 
      month == "mar" ~ "March",
      month == "apr" ~ "April",
      month == "may" ~ "May",
      month == "jun" ~ "June",
      month == "jul" ~ "July",
      month == "aug" ~ "August",
      month == "sep" ~ "September",
      month == "oct" ~ "October",
      month == "nov" ~ "November",
      month == "dec" ~ "December"
    )
  ) 
```

### Join the datasets

```{r}
merging_df =
  left_join(pols_month_df, snp_df, by = c("year", "month")) |> 
  left_join(unemployment_df, by = c("year", "month"))
```

### description of each dataset

For the pols-month dataset, the dimension is (`r dim(pols_month_df)`), the range of years is ( `r range(pols_month_df$year, na.rm = TRUE)`) and the names of key values include `r names(pols_month_df)`.

For the snp dataset, the dimension is (`r dim(snp_df)`), the range of years is ( `r range(snp_df$year, na.rm = TRUE)`) and the names of key values include `r names(snp_df)`.

For the unemployment dataset, the dimension is (`r dim(unemployment_df)`), the range of years is ( `r range(unemployment_df$year, na.rm = TRUE)`) and the names of key values include `r names(unemployment_df)`.

For the resulting dataset, the dimension is (`r dim(merging_df)`), the range of years is ( `r range(merging_df$year, na.rm = TRUE)`) and the names of key values include `r names(merging_df)`.
