---
title: "p8105_hw2_fx2212"
author: "Fenglin Xie"
date: "2025-09-29"
output: github_document
---
```{r}
library(tidyverse)
library(readxl)
library(knitr)
```

## Problem 1 

### clean the data in pols-month.csv

```{r}
pols_month_df = 
  read_csv("fivethirtyeight_datasets/pols-month.csv") |> 
  janitor::clean_names() |> 
  separate(
    col = mon,
    into = c("year", "month", "day")
  ) |> 
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day),
    month = month.name[month],
    president = case_when(
      prez_gop == 1 ~ "gop",
      prez_dem == 1 ~ "dem",
      TRUE ~ NA_character_
    )
  ) |>
  select(-prez_dem, -prez_gop, -day) 
  
```

### clean the data in snp.csv

```{r}
snp_df =
  read_csv("fivethirtyeight_datasets/snp.csv") |> 
  janitor::clean_names() |>
  mutate(
    date = mdy(date)
  ) |> 
  mutate(
    year = year(date),
    month = month(date),
    month = month.name[month]
  ) |> 
  select(year, month, close) |> 
  arrange(year, match(month, month.name))
```

### tidy the unemployment data

```{r}
unemployment_df = 
  read_csv("fivethirtyeight_datasets/unemployment.csv") |> 
  janitor::clean_names() |> 
  pivot_longer(
    cols = jan:dec,
    names_to = "month",
    values_to = "unemployment_rate"
  ) |> 
  mutate(
      month = case_when(
      month == "jan" ~ "January",
      month == "feb" ~ "February", 
      month == "mar" ~ "March",
      month == "apr" ~ "April",
      month == "may" ~ "May",
      month == "jun" ~ "June",
      month == "jul" ~ "July",
      month == "aug" ~ "August",
      month == "sep" ~ "September",
      month == "oct" ~ "October",
      month == "nov" ~ "November",
      month == "dec" ~ "December"
    )
  ) 
```

### Join the datasets

```{r}
merging_df =
  left_join(pols_month_df, snp_df, by = c("year", "month")) |> 
  left_join(unemployment_df, by = c("year", "month"))
```

### description of each dataset

For the pols-month dataset, the dimension is (`r dim(pols_month_df)`), the range of years is ( `r range(pols_month_df$year, na.rm = TRUE)`) and the names of key values include `r names(pols_month_df)`.

For the snp dataset, the dimension is (`r dim(snp_df)`), the range of years is ( `r range(snp_df$year, na.rm = TRUE)`) and the names of key values include `r names(snp_df)`.

For the unemployment dataset, the dimension is (`r dim(unemployment_df)`), the range of years is ( `r range(unemployment_df$year, na.rm = TRUE)`) and the names of key values include `r names(unemployment_df)`.

For the resulting dataset, the dimension is (`r dim(merging_df)`), the range of years is ( `r range(merging_df$year, na.rm = TRUE)`) and the names of key values include `r names(merging_df)`.

## Problem 2

### Read and clean the Mr. Trash Wheel sheet

```{r}
mr_trash_wheel =
  read_excel("202509 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel") |> 
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |> 
  mutate(
    sports_balls = as.integer(round(sports_balls)),
    wheel = "Mr. Trash Wheel",
    year = as.numeric(year)
  )
  
```

### Read and clean the Professor Trash Wheel

```{r}
prof_trash_wheel <- read_excel("202509 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel") |> 
  janitor::clean_names()  |> 
  filter(!is.na(dumpster))  |> 
  mutate(
    wheel = "Professor Trash Wheel",
    year = as.numeric(year)
  )

```

### Read and clean the Gwynns Falls Trash Wheel
```{r}
gwynns_trash_wheel <- read_excel("202509 Trash Wheel Collection Data.xlsx", sheet = "Gwynns Falls Trash Wheel")  |> 
  janitor::clean_names()  |> 
  filter(!is.na(dumpster))  |> 
  mutate(
    wheel = "Gwynns Falls Trash Wheel",
    year = as.numeric(year)
  )
```

### combine dataset

```{r}
all_trash_wheel <- bind_rows(mr_trash_wheel, prof_trash_wheel, gwynns_trash_wheel)
```

### description of the resulting dataset

```{r}
total_prof_weight =
  all_trash_wheel  |> 
  filter(wheel == "Professor Trash Wheel")  |> 
  summarise(total_weight = sum(weight_tons, na.rm = TRUE))
```


```{r}
gwynns_june_2022 =
  all_trash_wheel |> 
  filter(wheel == "Gwynns Falls Trash Wheel", 
         year == 2022, 
         month == "June")  |> 
  summarise(total_cigarettes = sum(cigarette_butts, na.rm = TRUE))
```


The number of observations in the resulting dataset is `r nrow(all_trash_wheel)` and the key variables include `r names(all_trash_wheel)`.

The total weight of trash collected by Professor Trash Wheel is `r total_prof_weight` and the total number of cigarette butts collected by Gwynns in June of 2022 is `r gwynns_june_2022`.

## Problem 3

### Create a single, well-organized dataset with all the information contained in these data files--"rent_data"

```{r}
zip_codes = read_csv("zillow_data/Zip Codes.csv") |> 
  janitor::clean_names() |> 
  select(county, zip_code, neighborhood) |> 
  mutate(
    borough = case_when(
      county == "Bronx" ~ "Bronx",
      county == "Kings" ~ "Brooklyn", 
      county == "New York" ~ "Manhattan",
      county == "Queens" ~ "Queens",
      county == "Richmond" ~ "Staten Island",
      TRUE ~ county
    ),
    zip_code = as.character(zip_code)
  ) |> 
  distinct(zip_code, .keep_all = TRUE) |> 
  filter(!is.na(neighborhood))

zillow_rent = read_csv("zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv", 
                       col_types = cols(RegionName = col_character()),
                       show_col_types = FALSE) |>
  filter(RegionType == "zip") |>
  select(RegionName, matches("^\\d{4}-\\d{2}-\\d{2}$")) |>
  pivot_longer(
    cols = -RegionName,
    names_to = "date",
    values_to = "rent_price"
  ) |> 
  mutate(
    date = as.Date(date),
    zip_code = RegionName
  ) |>
  select(zip_code, date, rent_price) |> 
  filter(!is.na(rent_price))

rent_data <- zillow_rent |> 
  inner_join(zip_codes, by = "zip_code")

```

The number of total observations is `r nrow(rent_data)`, the number of unique ZIP codes is `r n_distinct(rent_data$zip_code)`, and the number of unique neighborhoods is `r n_distinct(rent_data$neighborhood)`.

### Finding ZIP codes appear in the ZIP code dataset but not in the Zillow Rental Price dataset

```{r}
missing_in_zillow = zip_codes |> 
  anti_join(zillow_rent, by = "zip_code")
```

The number of ZIP codes appearing in the ZIP code dataset but not in the Zillow Rental Price dataset is `r nrow(missing_in_zillow)`. There might be some reasons:

*  May be commercial or industrial areas with no residential rental data
*  Might be newly developed areas not included during data collection
*  Data coverage may be incomplete or Zillow lacks sufficient data points



### comparing rental prices in January 2021 to prices in January 2020

```{r}
rent_jan_2020 = rent_data |> 
  filter(year(date) == 2020, month(date) == 1)

rent_jan_2021 = rent_data  |> 
  filter(year(date) == 2021, month(date) == 1)

rent_change = rent_jan_2020  |> 
  select(zip_code, borough, neighborhood, rent_2020 = rent_price)  |> 
  inner_join(
    rent_jan_2021  |>  
      select(zip_code, rent_2021 = rent_price),
    by = "zip_code"
  )  |> 
  mutate(
    rent_change = rent_2021 - rent_2020,
    percent_change = (rent_2021 - rent_2020) / rent_2020 * 100
  )

top_drops = rent_change  |> 
  arrange(percent_change)  |> 
  head(10)

top_drops_table = top_drops  |> 
  select(
    ZIP_Code = zip_code,
    Borough = borough,
    Neighborhood = neighborhood,
    `Rent_Jan_2020` = rent_2020,
    `Rent_Jan_2021` = rent_2021,
    `Price_Drop` = rent_change,
    `Percent_Drop` = percent_change
  ) |> 
  mutate(
    across(c(`Rent_Jan_2020`, `Rent_Jan_2021`, `Price_Drop`), ~round(., 2)),
    `Percent_Drop` = round(`Percent_Drop`, 2)
  )

kable(top_drops_table, 
      caption = "Top 10 ZIP Codes with Largest Rent Drops from Jan 2020 to Jan 2021")

```


### comments

```{r}
borough_distribution = top_drops  |> 
    count(Borough = borough)  |> 
    arrange(desc(n))
```

We can find that Manhattan has the most ZIP codes in the top drops list (`r borough_distribution$n` ZIP codes), in fact, it completes the top 10, which may indicates that the COVID-19 impact: These significant rent drops likely reflect the substantial impact of the COVID-19 pandemic on New York City's rental market, particularly in high-density urban areas and communities reliant on commuters. The widespread adoption of remote work led many residents to move away from city centers, seeking more spacious and affordable housing options.
