---
title: "p8105_hw2_fx2212"
author: "Fenglin Xie"
date: "2025-09-29"
output: github_document
---
```{r}
library(tidyverse)
library(readxl)
```

## Problem 1 

### clean the data in pols-month.csv

```{r}
pols_month_df = 
  read_csv("fivethirtyeight_datasets/pols-month.csv") |> 
  janitor::clean_names() |> 
  separate(
    col = mon,
    into = c("year", "month", "day")
  ) |> 
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day),
    month = month.name[month],
    president = case_when(
      prez_gop == 1 ~ "gop",
      prez_dem == 1 ~ "dem",
      TRUE ~ NA_character_
    )
  ) |>
  select(-prez_dem, -prez_gop, -day) 
  
```

### clean the data in snp.csv

```{r}
snp_df =
  read_csv("fivethirtyeight_datasets/snp.csv") |> 
  janitor::clean_names() |>
  mutate(
    date = mdy(date)
  ) |> 
  mutate(
    year = year(date),
    month = month(date),
    month = month.name[month]
  ) |> 
  select(year, month, close) |> 
  arrange(year, match(month, month.name))
```

### tidy the unemployment data

```{r}
unemployment_df = 
  read_csv("fivethirtyeight_datasets/unemployment.csv") |> 
  janitor::clean_names() |> 
  pivot_longer(
    cols = jan:dec,
    names_to = "month",
    values_to = "unemployment_rate"
  ) |> 
  mutate(
      month = case_when(
      month == "jan" ~ "January",
      month == "feb" ~ "February", 
      month == "mar" ~ "March",
      month == "apr" ~ "April",
      month == "may" ~ "May",
      month == "jun" ~ "June",
      month == "jul" ~ "July",
      month == "aug" ~ "August",
      month == "sep" ~ "September",
      month == "oct" ~ "October",
      month == "nov" ~ "November",
      month == "dec" ~ "December"
    )
  ) 
```

### Join the datasets

```{r}
merging_df =
  left_join(pols_month_df, snp_df, by = c("year", "month")) |> 
  left_join(unemployment_df, by = c("year", "month"))
```

### description of each dataset

For the pols-month dataset, the dimension is (`r dim(pols_month_df)`), the range of years is ( `r range(pols_month_df$year, na.rm = TRUE)`) and the names of key values include `r names(pols_month_df)`.

For the snp dataset, the dimension is (`r dim(snp_df)`), the range of years is ( `r range(snp_df$year, na.rm = TRUE)`) and the names of key values include `r names(snp_df)`.

For the unemployment dataset, the dimension is (`r dim(unemployment_df)`), the range of years is ( `r range(unemployment_df$year, na.rm = TRUE)`) and the names of key values include `r names(unemployment_df)`.

For the resulting dataset, the dimension is (`r dim(merging_df)`), the range of years is ( `r range(merging_df$year, na.rm = TRUE)`) and the names of key values include `r names(merging_df)`.

## Problem 2

### Read and clean the Mr. Trash Wheel sheet

```{r}
mr_trash_wheel =
  read_excel("202509 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel") |> 
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |> 
  mutate(
    sports_balls = as.integer(round(sports_balls)),
    wheel = "Mr. Trash Wheel",
    year = as.numeric(year)
  )
  
```

### Read and clean the Professor Trash Wheel

```{r}
prof_trash_wheel <- read_excel("202509 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel") |> 
  janitor::clean_names()  |> 
  filter(!is.na(dumpster))  |> 
  mutate(
    wheel = "Professor Trash Wheel",
    year = as.numeric(year)
  )

```

### Read and clean the Gwynnda Trash Wheel
```{r}
gwynns_trash_wheel <- read_excel("202509 Trash Wheel Collection Data.xlsx", sheet = "Gwynns Falls Trash Wheel")  |> 
  janitor::clean_names()  |> 
  filter(!is.na(dumpster))  |> 
  mutate(
    wheel = "Gwynns Falls Trash Wheel",
    year = as.numeric(year)
  )
```

### combine dataset

```{r}
all_trash_wheel <- bind_rows(mr_trash_wheel, prof_trash_wheel, gwynns_trash_wheel)
```

### description of the resulting dataset

```{r}
total_prof_weight =
  all_trash_wheel  |> 
  filter(wheel == "Professor Trash Wheel")  |> 
  summarise(total_weight = sum(weight_tons, na.rm = TRUE))
```


```{r}
gwynns_june_2022 =
  all_trash_wheel |> 
  filter(wheel == "Gwynns Falls Trash Wheel", 
         year == 2022, 
         month == "June")  |> 
  summarise(total_cigarettes = sum(cigarette_butts, na.rm = TRUE))
```


The number of observations in the resulting dataset is `r nrow(all_trash_wheel)` and the key variables include `r names(all_trash_wheel)`.

The total weight of trash collected by Professor Trash Wheel is `r total_prof_weight` and the total number of cigarette butts collected by Gwynnda in June of 2022 is `r gwynns_june_2022`.

## Problem 3

### Create a single, well-organized dataset with all the information contained in these data files

```{r}
zip_codes = read_csv("zillow_data/Zip Codes.csv")
Zip_zori_uc_sfrcondomfr_sm_month_NYC = read_csv("zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv")


```

